Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex
small fixs

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Secure Composition of Robust and Optimising Compilers}

@@

\gls{ms} is a security property of programs that, e.g., ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile context, i.e., where it interacts with other possibly malicious programs.
Secure compilers can be described formally with the theory on {\em robust preservation}~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
Unfortunately, it is unclear if and how robust preservation supports modularity.
Modularity is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
Compilation passes are typically developed as a standalone module that can plugged into a compilation pipeline.
Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfills a property of interest.
Without a modular framework, there is no hope to enable this separation of concerns, where, e.g., a secure pass that ensures \gls{ms} can be developed independently of a secure pass for \gls{cct}.
%% MK: no idea how/where to fit this in or if this is even necessary
%Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
The aim of this paper is to devise the theoretical foundation to be able to modularily build a secure compiler that is capable of dealing with the following example.
%\MPin{ MK: hope it's solved now
%	this does not really introduce the compositionality issue.
%	i think the problem is the last sentence.
%	try and divert the story after the 'because' line, to go into the need for composing compilers
%}
\begin{example}[strncpy]\label{ex:strncpy}
Consider the program below which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Secure Composition of Robust and Optimising Compilers}
% Composing Secure and Optimising Compilers
%

@@

\gls{ms} is a security property of programs that, e.g., ensures that there are no out-of-bounds accesses as well as no reads or writes to uninitialized memory.
Without memory safety, private data can easily leak or be corrupted~\cite{lemay2021ccc}.
Because of this, \gls{ms} is a requirement for programs to attain an even stronger security property, namely \gls{cct}.
Programs that attain \gls{cct} are secure with respect to timing attacks, and this is the gold standard for secure applications, including cryptographic ones since timing attacks can expose secrets via a simple performance analysis~\cite{kocher1996timing}.
Timing attacks have become more and more relevant recently, especially given the era of cloud-computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, the literature shows that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
This is of course not what programmers expect and necessitates the need for a {\em secure} compiler that ensures such mitigations persist in the code they generate.
More generally, secure compilers should preserve security properties even when the program runs in a hostile context such as the cloud one, i.e., where it interacts with other possibly malicious programs.
An example of such secure compilers are those that ascribe to the theory of robust compilation~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}, which takes this interaction between components and potentially malicious contexts into account.
Unfortunately, it is unclear if and how robust compilation supports compositionality.

% Compositionality is an inherent part of software engineering and real-world compilers are no exception to that~\cite{lattner2004llvm,googlev8}.
Compilation passes are typically developed as a standalone module to be plugged into a compilation pipeline and composed with other passes.
Different kinds of such passes exist, e.g., there are usually several compilation passes that perform optimisations~\cite{lattner2004llvm,googlev8,androidstudio,kuepper2023cryptopt,manjikian1997fusion,wegman1991ccp} or source-code instrumentations~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact} that enforce that the compiled program fulfils a property of interest.
Without a modular framework, there is no hope to enable this separation of concerns, where, e.g., a secure pass that ensures \gls{ms} can be developed independently of a secure pass for \gls{cct}.
%% MK: no idea how/where to fit this in or if this is even necessary
%Because of the potentially hostile context, secure compilers attain security by relying on protection mechanisms, such as ... and ...\MK{cits}.
The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler as the composition of several smaller ones.
Let us now introduce a minimal example of what such a secure compiler should be capable of dealing with.
%\MPin{ MK: hope it's solved now
%	this does not really introduce the compositionality issue.
%	i think the problem is the last sentence.
%	try and divert the story after the 'because' line, to go into the need for composing compilers
%}
\begin{example}[strncpy]\label{ex:strncpy}
Consider the program below which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:
